The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-09-06 analysis/paper/paper.qmd

# Concatenate ngrams
emu$text_clean <- emu |> 
  c_ngrams(n_max = 3, text_col = "text_clean",
           title_col = "title", min_freq = 20, lemma = TRUE)

@@

# TODO Replace title column with ID column

# Unnest tokens and remove words of one and two letters
emu_words <- emu |> 
  select(title, text_clean) |> 
  unnest_tokens(output = word, input = text_clean) |>  # remove punctuation, convert to lower-case, separate all words
  anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
  mutate(word = textstem::lemmatize_words(word)) |>  # lemmatise words
  filter(nchar(word) >= 3)  # keep only words that are at least three letters long

@@

# Prepare document-term matrix as input for LDA model
# Reduce the size of the DTM by excluding words with a frequency lower than 5
emu_dtm <- convert_to_dtm(emu_words, 
                          title_col = "title", 
                          word_col = "word", 
                          min_term_freq = 5)

@@

# TODO change to pivot_longer
vizDataFrameWithYear <- melt(cbind(data.frame(topic_prop_per_year), 
                           document = factor(rownames(topic_prop_per_year))), 
                     variable.name = "topic", id.vars = "document") |> 
    left_join(select(emu_theses, title, "year" = graduation_year), by = c("document" = "title"))

@@

emu_documents |> 
  inner_join(emu |> 
               select(title, graduation_year), 
             by = c("document" = "title")) |> 
  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  facet_wrap(~ graduation_year, ncol = 1, scales = "free", drop = TRUE)

emu_documents |> 
  inner_join(emu |> 
               select(title, graduation_year), 
             by = c("document" = "title")) |> 
  ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
  geom_col()
```

@@

emu_bigrams <- emu |> 
  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)

@@

bigram_tf_idf <- emu_bigrams |> 
  count(title, ngram) |> 
  bind_tf_idf(ngram, title, n) |> 
  arrange(desc(tf_idf))

@@

emu_trigrams <- emu |> 
  get_ngrams(n = 3, title_col = "title", text_col = "text_raw")

@@

trigram_tf_idf <- emu_trigrams |> 
  count(title, ngram) |> 
  bind_tf_idf(ngram, title, n) |> 
  arrange(desc(tf_idf))

@@

emu_bigrams_top <- emu_bigrams |>
  group_by(title) |>
  count(ngram, sort = TRUE) |>
  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  ungroup() |>
  mutate(bigram = reorder(ngram, n))

@@

emu_bigrams_top |>
  filter(title == unique(title)[1:10]) |> 
  ggplot(aes(x = ngram, y = n, fill = title)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ title, scales = "free", ncol = 2) +
  theme(legend.position = "none") 
```

@@

## Calculate tf-idf statistic
emu_tf_idf <- emu_words |> 
  get_tf_idf(title_col = "title", word_col = "word")

## Get top n words most specific to each thesis 
emu_tf_idf |> 
  group_by(title) |> 
  top_n(5, tf_idf) |> 
  arrange(desc(tf_idf), .by_group = TRUE)

@@

# Preview term frequency distributions in a subset of theses
emu_words_count <- emu_words |> 
  count(title, word, sort = TRUE) |> 
  ungroup()

emu_words_total <- emu_words_count |> 
  group_by(title) |> 
  summarise(total = sum(n))

@@

ggplot(emu_words_count |> 
         filter(title == unique(title)[1:6]),
       aes(n/total, fill = title)) +
  geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
  xlim(NA, 0.0009) +
  facet_wrap(~ title, ncol = 3, scales = "free_y")
```

```{r}
# Examine Zipf's law for the EMU theses
## Calculate rank and term frequency
freq_by_rank <- emu_words_count |> 
  group_by(title) |> 
  mutate(rank = row_number(),
         `term frequency` = n/total)

## Visualise Zipf's law
freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = title)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  scale_x_log10() +
  scale_y_log10()

@@

freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = title)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  geom_abline(intercept = -1.44, slope = -0.7, color = "gray50", linetype = 2) +
  scale_x_log10() +
  scale_y_log10()
```

@@

## Visualise the top n words with the highest tf-idf values for each of a subset of theses
emu_words_count |> 
  get_tf_idf(title_col = "title", word_col = "word") |> 
  filter(title == unique(title)[1:2]) |>  # Keep only the first 6 theses
  arrange(desc(tf_idf)) |> 
  mutate(word = factor(word, levels = rev(unique(word)))) |> 
  group_by(title) |> 
  top_n(15) |> 
  ungroup() |> 
  ggplot(aes(word, tf_idf, fill = title)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~ title,  ncol = 3, scales = "free") +
  coord_flip()
```

@@

# To what extent do words co-occur with other words?
tidy_pmi <- get_pmi(emu, title_col = "title", text_col = "text_clean")
```

@@

tidy_words <- emu |> 
  select(title, text_clean) |> 
  unnest_tokens(word, text_clean) |> 
  add_count(word) |> 
  filter(n >= 50) |> 
  select(-n)

# Summarise word embeddings into document embeddings for modeling purposes
word_matrix <- tidy_words |> 
  count(title, word) |> 
  cast_sparse(title, word, n)

>>>>

# Concatenate ngrams
emu$text_clean <- emu |> 
  c_ngrams(n_max = 3, text_col = "text_clean",
           id_col = "ID", min_freq = 20, lemma = TRUE)

@@

# Unnest tokens and remove words of one and two letters
emu_words <- emu |> 
  select(ID, text_clean) |> 
  unnest_tokens(output = word, input = text_clean) |>  # remove punctuation, convert to lower-case, separate all words
  anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
  mutate(word = textstem::lemmatize_words(word)) |>  # lemmatise words
  filter(nchar(word) >= 3)  # keep only words that are at least three letters long

@@

# Prepare document-term matrix as input for LDA model
# Reduce the size of the DTM by excluding words with a frequency lower than 5
emu_dtm <- convert_to_dtm(emu_words, 
                          id_col = "ID", 
                          word_col = "word", 
                          min_term_freq = 5)

@@

# TODO change to pivot_longer
vizDataFrameWithYear <- melt(cbind(data.frame(topic_prop_per_year), 
                           document = factor(rownames(topic_prop_per_year))), 
                     variable.name = "topic", id.vars = "document") |> 
    left_join(select(emu_theses, ID, "year" = graduation_year), by = c("document" = "ID"))

@@

emu_documents |> 
  inner_join(emu |> 
               select(ID, graduation_year), 
             by = c("document" = "ID")) |> 
  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  facet_wrap(~ graduation_year, ncol = 1, scales = "free", drop = TRUE)

emu_documents |> 
  inner_join(emu |> 
               select(ID, graduation_year), 
             by = c("document" = "ID")) |> 
  ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
  geom_col()
```

@@

emu_bigrams <- emu |> 
  get_ngrams(n = 2, id_col = "ID", text_col = "text_clean", stem = TRUE)

@@

bigram_tf_idf <- emu_bigrams |> 
  count(ID, ngram) |> 
  bind_tf_idf(ngram, ID, n) |> 
  arrange(desc(tf_idf))

@@

emu_trigrams <- emu |> 
  get_ngrams(n = 3, id_col = "ID", text_col = "text_raw")

@@

trigram_tf_idf <- emu_trigrams |> 
  count(ID, ngram) |> 
  bind_tf_idf(ngram, ID, n) |> 
  arrange(desc(tf_idf))

@@

emu_bigrams_top <- emu_bigrams |>
  group_by(ID) |>
  count(ngram, sort = TRUE) |>
  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  ungroup() |>
  mutate(bigram = reorder(ngram, n))

@@

emu_bigrams_top |>
  filter(ID == unique(ID)[1:10]) |> 
  ggplot(aes(x = ngram, y = n, fill = ID)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ ID, scales = "free", ncol = 2) +
  theme(legend.position = "none") 
```

@@

## Calculate tf-idf statistic
emu_tf_idf <- emu_words |> 
  get_tf_idf(id_col = "ID", word_col = "word")

## Get top n words most specific to each thesis 
emu_tf_idf |> 
  group_by(ID) |> 
  top_n(5, tf_idf) |> 
  arrange(desc(tf_idf), .by_group = TRUE)

@@

# Preview term frequency distributions in a subset of theses
emu_words_count <- emu_words |> 
  count(ID, word, sort = TRUE) |> 
  ungroup()

emu_words_total <- emu_words_count |> 
  group_by(ID) |> 
  summarise(total = sum(n))

@@

ggplot(emu_words_count |> 
         filter(ID == unique(ID)[1:6]),
       aes(n/total, fill = ID)) +
  geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
  xlim(NA, 0.0009) +
  facet_wrap(~ ID, ncol = 3, scales = "free_y")
```

```{r}
# Examine Zipf's law for the EMU theses
## Calculate rank and term frequency
freq_by_rank <- emu_words_count |> 
  group_by(ID) |> 
  mutate(rank = row_number(),
         `term frequency` = n/total)

## Visualise Zipf's law
freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = ID)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  scale_x_log10() +
  scale_y_log10()

@@

freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = ID)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  geom_abline(intercept = -1.44, slope = -0.7, color = "gray50", linetype = 2) +
  scale_x_log10() +
  scale_y_log10()
```

@@

## Visualise the top n words with the highest tf-idf values for each of a subset of theses
emu_words_count |> 
  get_tf_idf(id_col = "ID", word_col = "word") |> 
  filter(ID == unique(ID)[1:2]) |>  # Keep only the first 6 theses
  arrange(desc(tf_idf)) |> 
  mutate(word = factor(word, levels = rev(unique(word)))) |> 
  group_by(ID) |> 
  top_n(15) |> 
  ungroup() |> 
  ggplot(aes(word, tf_idf, fill = ID)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~ ID,  ncol = 3, scales = "free") +
  coord_flip()
```

@@

# To what extent do words co-occur with other words?
tidy_pmi <- get_pmi(emu, id_col = "ID", text_col = "text_clean")
```

@@

tidy_words <- emu |> 
  select(ID, text_clean) |> 
  unnest_tokens(word, text_clean) |> 
  add_count(word) |> 
  filter(n >= 50) |> 
  select(-n)

# Summarise word embeddings into document embeddings for modeling purposes
word_matrix <- tidy_words |> 
  count(ID, word) |> 
  cast_sparse(ID, word, n)
